---
title: 'landscapemetrics'
author: Jakub Nowosad
date: '2019-03-23'
slug: landscapemetrics-bp
categories: []
tags:
  - spatial
  - geocompr
  - sil
  - rstats
  - landscape-ecology
draft: yes
---

# Data preparation

Every software for calculations of landscape metrics, such as FRAGSTATS or **landscapemetrics** expects the input data to be in a projected coordinate reference system instead of a geographic coordinate reference system. 
This is due to a fact that geographic coordinate reference systems are expressed in degrees, and one degree on the equator has a different length in meters than one degree on a middle latitude.
Projected coordinate reference systems, on the other hand, have a linear unit of measurement such as meters.
This allows for proper calculations of distances or areas.

In this case, the input data (`studyarea.tif` and `jb.csv`) is in a geographic coordinate reference system called WGS84.
Therefore, it is important to reproject the data before calculating any landscape metrics.
Reprojection of raster objects is possible with an R package called **gdalUtils**:

```{r, eval=FALSE}
library(gdalUtils) 
library(sf)
# unzip the file
unzip("data/cci_lc_australia.zip", "data/cci_lc_australia.tif")
# reprojecting the raster data to a local projected coordinate reference system
# read more at http://spatialreference.org/ref/sr-org/7564/
gdalwarp(srcfile = "data/cci_lc_australia.tif",
         dstfile = "data/cci_lc_australia_aea.vrt",
         of = "VRT",
         t_srs = "+proj=aea +lat_1=-18 +lat_2=-36 +lat_0=0 +lon_0=132 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs")
# compressing the output file
gdal_translate(src_dataset = "data/cci_lc_australia_aea.vrt",
               dst_dataset = "data/cci_lc_australia_aea.tif",
               co = "COMPRESS=LZW")
```

Reading and reprojecting vector (points) objects could be done with the **sf** package:

```{r, eval=FALSE}
# reading the vector data and reprojecting it
points = read.csv("data/sample_points.csv") %>% 
  st_as_sf(coords = c("longitude", "latitude")) %>% 
  st_set_crs(4326) %>% 
  st_transform("+proj=aea +lat_1=-18 +lat_2=-36 +lat_0=0 +lon_0=132 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs ")

# saving the new object to a spatial data format GPKG
st_write(points, "data/sample_points.gpkg")
```

# Simple calculations of landscape metrics

Now, we can read the new files - the new raster with the `raster()` function and the new points file with the `st_read()` function:

```{r, message=FALSE}
# attaching the packages used in the next examples
library(raster)
library(sf)
library(landscapemetrics)
library(purrr)
# reading the raster and vector datasets
r = raster("data/cci_lc_australia_aea.tif")
points = st_read("data/sample_points.gpkg")
```

The **landscapemetrics** package has a function designed for calculations of landscape metrics for a given buffer.
It is called `sample_lsm()` and it expects, at least three inputs - a raster, a vector (points), and a buffer size.
As a default, it calculates all of the available metrics:

```{r, warning=FALSE}
square_all  = sample_lsm(lc, 
                         points = sample_points,
                         size = 30000)
```

<!--adds info about warnings-->
This function also allows for calculating a selected level (`"landscape"`, `"class"`, or `"patch"`):

```{r, warning=FALSE}
square_landscape = sample_lsm(lc,
                                   points = sample_points,
                                   size = 30000,
                                   evel = "landscape")
square_landscape
```

... or just a single metric, for example SHDI on a landscape level (`"lsm_l_shdi"`):

```{r, warning=FALSE}
square_shdi = sample_lsm(lc,
                                   points = sample_points,
                                   size = 30000,
                              what = "lsm_l_shdi") 
square_shdi
```

As a default, it uses a square buffer, but it could be changed to a circle one.

You can read more about this function in [its help file](https://r-spatialecology.github.io/landscapemetrics/reference/sample_lsm.html) and about the structure of the output on [the package website](https://r-spatialecology.github.io/landscapemetrics/index.html).

# Calculations of landscape metrics for many buffers

Finally, it is also possible to calculate landscapemetrics on different buffer sizes.
You need to select what sizes you are interested in, and next use the `sample_lsm()` function inside a `map()` function.

```{r, warning=FALSE}
# create buffers of 1000 and 2000 meters 
sizes = c(1000, 2000)
# calculate landscape metrics for each buffer
two_sizes_output = sizes %>% 
  map(~sample_lsm(lc, points = sample_points size = .))
```

It calculates all of the selected metrics for all of the buffers.
The result is a list, where each element of the output list describes each buffer.

```{r}
# select the result for the first buffer
two_sizes_output[[1]]
# select the result for the second buffer
two_sizes_output[[2]]
```

# Summary

